Polynomial phase estimation by least squares phase unwrapping

### 2013-12-16 paper.tex
minor
minor?
added product CPF to simulation software and ran some simulation with it. It does have a lower threshold than the other estimators

\author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
    \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.  Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.   Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.  B. Moran is with the Defense Science Institute, University of Melbourne, Vic. 3010, Australia.}}
% The paper headers 
\markboth{Polynomial phase estimation by phase unwrapping}{DRAFT \today}

@@

@@

A \term{lattice},  $\Lambda$, is a discrete subset of points in $\reals^n$ for which
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^d \}
\]
where $\Bbf \in \reals^{n \times d}$ has rank $d$ and is called the generator matrix.  If $n = d$ the lattice is said to be of full rank.  Lattices are discrete Abelian groups under vector addition.  They are subgroups of the Euclidean group $\reals^n$.  Lattices naturally give rise to tessellations of $\reals^n$ by the specification of a set of coset representatives for the quotient $\reals^n / \Lambda$.  One choice for a set of coset representatives is a fundamental parallelepiped, the parallelepiped generated by the columns of a generator matrix.  Another choice is based on the Voronoi cell, those points from $\reals^n$ nearest (with respect to the Euclidean norm in this paper) to the lattice point at the origin.  It is always possible to construct a rectangular set of representatives, as the next proposition will show.  These rectangular regions are used to describe the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.  These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator.
% in Section~\ref{sec:least-squar-unwr}.

@@

Given a polynomial $g(x) = a_0 + a_1x + \dots + a_m x^m$, let
\[
\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m \end{array} \right]^\prime,
\]
denote the column vector of length $m+1$ containing the coefficients of $g$, where $^\prime$ indicates transpose.  If $y$ and $z$ differ by a polynomial in $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and then also $\coef(y) = \coef(z) + \coef(p)$.
%\footnote{In group theory terminology $\coef(\cdot)$ coupled with vector addition is called a \term{group homomorphism}.\index{group homomorphism}}. 
  Consider the set
\[
L_{m+1} = \{ \coef(p) \mid p \in \mathcal{Z} \}
\]
containing the coefficient vectors corresponding to the polynomials in $\mathcal{Z}$.  Since the integer valued polynomials form a basis for $\mathcal{Z}$,
\begin{align*}
L_{m+1} &= \{ \coef(c_0 p_0 + c_1p_1 + \dots + c_mp_m) \mid c_i \in \ints \} \\
&= \{ c_0 \coef(p_0) + \dots + c_m\coef(p_m) \mid c_i \in \ints \}.
\end{align*}
Let
\[
\Pbf = \left[ \begin{array}{cccc} \coef(p_0)& \coef(p_1)& \dots& \coef(p_m)  \end{array} \right]
\]
be the $m+1$ by $m+1$ matrix with columns given by the coefficients of the integer valued polynomials.  Then,
\[
L_{m+1} = \{ \xbf = \Pbf\ubf \mid \ubf \in \ints^{m+1} \}
\]
and it is clear that $L_{m+1}$ is an $m+1$ dimensional lattice.  That is, the set of coefficients of the polynomials from $\mathcal{Z}$ forms a lattice with generator matrix $\Pbf$. We can restate Lemma~\ref{lem:circpolysampledthm} as:
\begin{corollary}\label{cor:circpolysampledcoef}
Two polynomial phase signals $s(t) = e^{2\pi j y(t)}$  and $r(t) = e^{2\pi j z(t)}$ are aliases if and only if $\coef(y)$ and $\coef(z)$ differ by a lattice point in $L_{m+1}$.
\end{corollary}

@@

This minimisation problem can be posed as a nearest lattice point problem~\cite{McKilliam2009asilomar_polyest_lattice, McKilliam2010thesis}.  Write $SS$ as
\[
SS(\mubf) = \sum_{n=1}^{N}{ \left( \Theta_{n} - W_n - \sum_{k = 0}^{m}{\mu_k n^k} \right)^2 },
\]
where $W_n = \round{\Theta_{n} - \sum_{k = 0}^{m}{\mu_k n^k} }$ for $n = 1, \dots, N$ are integers called \emph{wrapping variables}.  If we consider $W_1,\dots,W_N$ as nuisance parameters to be estimated then $SS$ can be considered as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$, the minimiser with respect to $W_1,\dots,W_N$ being $SS(\mubf)$.  The LSU estimator can be found by jointly minimising over $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.   This joint minimisation problem can be solved by computing a nearest point in a lattice.  To see this write $SS$ as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$,
\begin{equation}\label{eq:SS_in_vector_form_with_u_nuisance}
SS(\mubf, \wbf) = \| \thetabf - \Xbf\mubf - \wbf \|^2,
\end{equation}
where $\|\cdot\|^2$ denotes the squared Euclidean norm, $\Xbf$ is the rectangular Vandermonde matrix
\begin{equation}\label{eq:Xvandermonde}
\Xbf = \left[ \begin{array}{ccccc} 
1 & 1 & 1 & \cdots & 1 \\
1 & 2 & 4 & \cdots & 2^m \\
\vdots & \vdots  & \vdots & \ddots & \vdots \\
1 & N & N^2 & \cdots & N^m \\
\end{array} \right],
\end{equation}
and the column vectors $\wbf = [W_1, \dots, W_N]^\prime$ and $\thetabf = [\Theta_1, \dots, \Theta_N]^\prime$.   For fixed $\wbf$ the minimiser of $SS(\mubf, \wbf)$ with respect to $\mubf$ is
\begin{equation} \label{eq:hat_theta_cal}
\mubf(\wbf) = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime( \thetabf - \wbf) = \Xbf^+( \thetabf - \wbf),
\end{equation}
where $\Xbf^+ = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime$ is the Moore-Penrose pseudoinverse of $\Xbf$.  We may thus compute $\widehat{\mubf}$ as $\widehat{\mubf} = \mu(\widehat{\wbf})$, where $\widehat{\wbf}$ is the minimiser of
\[
SS(\wbf) = SS(\mubf(\wbf),\wbf) = \| \Qbf\thetabf - \Qbf\wbf \|^2,
\]
where $\Qbf = \Ibf - \Xbf\Xbf^+$ and $\Ibf$ is the $N \times N$ identity matrix\footnote{We have slightly abused notation here by reusing $SS$. This should not cause any confusion as $SS(\mubf)$ and $SS(\mubf,\wbf)$ and $SS(\wbf)$ are easily told apart by their inputs.}.

@@

The theorem makes statements about the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly on the difference $\tilde{\mubf} - \widehat{\mubf}$.   To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.  However, the two phases are obviously close, since the phases $\pm 0.5$ are actually the same.  In this case 
\[
\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01.
\] 
The same reasoning holds for $m > 0$.

@@

The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.  The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies that $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$. In this exceptional case other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and for this reason we have omitted them. %Distributions that do not satisfy these requirements are unlikely to be needed in practice.

@@

Tables~\ref{tablm1},~\ref{tablm3} and~\ref{tablm5} show the values of SNR at which the estimators reach a threshold for polynomial phase signals of order $m=1,3$ and $5$.  The values in the tables are approximated using Monte-Carlo simulation with $X_1, \dots, X_N$ pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian.  When $m=3$ the true coefficients are as in Figure~\ref{plot:m3small} and when $m=5$ the true coefficients are as in Figure~\ref{plot:m5} so that the coefficients are inside the range acceptable for the HAF, CPF and CPF-HAF estimators.  When $m=1$ (single frequency estimation) the LSU estimator can be computed exactly using the algorithm from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  In this case the value at which the LSU estimator reaches a threshold is similar to the maximum likelihood periodogram estimator~\cite{Quinn2001,Hannan1973,Rife1974}.  This property was observed in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  Both the LSU and periodogram estimators are strongly consistent, and so, the threshold values can be made arbitrarily small by choosing $N$ large.  This is indicated in the bottom row of Table~\ref{tablm1} by the $-\infty$ symbols.  When $m>1$ the LSU estimator can be exactly computed using the sphere decoder~\cite{Agrell2002,Viterbo_sphere_decoder_1999} in a reasonable amount of time when $N < 60$.  When $N > 60$ we use the approximate $K$-best algorithm.  We observe that the $K$-best algorithm produces estimates with the same threshold as the exactly computed LSU estimator when $N = 19$ and $49$.  As $N\rightarrow\infty$ the asymptotic threshold of the HAF estimator has been studied in~\cite{Porat_asympt_HAF_DPT_1996} and these values are indicated in the bottom rows of Tables~\ref{tablm3} and~\ref{tablm5}.  By Theorem~\ref{thm:asymp_proof} the LSU estimator is strongly consistent and the threshold can be made as small as desired by choosing $N$ large.  Whether the LSU estimator computed approximately using the $K$-best algorithm is strongly consistent is an open theoretical question.  That the $K$-best algorithm provides a close approximation of the exact LSU estimator for small $N$ is an encouraging observation.

\begin{table}[t]
\centering
\begin{tabular}{lrr}
 & LSU & ML  \\ \toprule
$N=16$ & 5 & 2  \\ 
$N=64$ & -2 & -4 \\ 
$N=256$ & -6 & -8 \\ 
$N=1024$ & -14 & -14 \\ 
$N\to\infty$ & $-\infty$ & $-\infty$  \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=1$}
\label{tablm1}
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{lrrrrr}
 & LSU & LSU ($K$-best) & CPF & HAF & ZW \\ \toprule
$N=19$ & 10 & 10 & 8 & 11 & 31 \\ 
$N=49$ & 7 & 7 & 5 & 9 & 29 \\ 
$N=99$ & & 6 & 3 & 7 & 29 \\ 
$N=199$ & & 3 & 1 & 5 & 28 \\ 
$N=299$ & & 1 & -1 & 5 & 27 \\ 
$N\to\infty$ & $-\infty$ & & & 3.77 & \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=3$}
\label{tablm3}
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{lrrrrr}
 & LSU & LSU ($K$-best) & CPF-HAF & HAF & ZW \\ \toprule
$N=19$ & 22  & 22 & 17 & 23 & $>100$ \\ 
$N=49$ & 17 & 17 & 14 & 21 & $>100$  \\ 
$N=99$ &  & 15 &  12 &  19 & $>100$ \\ 
$N=199$ & & 12 & 10 & 18 & $>100$ \\ 
$N=299$ &  & 11 & 9 & 17 & $>100$ \\ 
$N\to\infty$ & $-\infty$ &  &  & 16.6 &  \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=5$}
\label{tablm5}
\end{table}

@@

We have now shown that the highest order coefficient $\lambda_{m,N}$ converges as required. The remaining coefficients will be shown to converge by induction.  Assume that $N^k \lambda_{k,N} \rightarrow 0$ for all $k=r+1, r+2, \dots, m$, that is, assume that the $m-r$ highest order coefficients all converge as required. Let
\[
z_{N,r}(n) = \lambda_{0,N} + \lambda_{1,N} n + \dots + \lambda_{r,N} n^r.
\]
Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma_N(n)$ where 
\[
\sup_{n\in\{1,\dots,N\}}\abs{\gamma_N(n)} \rightarrow 0 \qquad \text{as $N\rightarrow\infty$}. 
\]
Now the bound from \eqref{eq:Deltazfracbound}, but applied using the $r$th difference, gives
 \begin{equation}\label{eq:zrbound}
\begin{split}
\left|\fracpart{  \Delta_h^r z_N(n')}\right| &= \left|\fracpart{  \Delta_h^r\gamma_N(n') + \Delta_h^r z_r(n') }\right| 
\\ &= |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
\end{split}
 \end{equation}
 where
\[
\epsilon = \Delta_h^r \gamma_N(n') \leq 2^r \sup_{n\in\{1,\dots,N\}}\abs{\gamma_N(n)} \rightarrow 0
\] 
as $N\rightarrow\infty$.  Choose $\delta$ and $\epsilon$ such that $2^r\delta < \tfrac{1}{4}$ and $|\epsilon| < \tfrac{1}{4}$.  Then, from \eqref{eq:zrbound} and from Lemma~\ref{lem:fracpartinternalsumlessdelta},
\[
\abs{\fracpart{ h^r r! \lambda_{r,N} }} < 2^r\delta + \abs{\epsilon}
\]
for all $h$ such that $1 \leq h \leq \tfrac{N}{2m}$.  Choosing $2^r\delta + |\epsilon| < 2^{-2r-1}$ and using the same iterative process as for the highest order coefficient $\lambda_{m,N}$  (see~\eqref{eq:startiterativearg}~to~\eqref{eq:enditerativearg}) we find that $N^r \lambda_{r,N} \rightarrow 0$ as $N\rightarrow\infty$.  The proof now follows by induction.
 \end{proof}

>>>>

\author{Robby~G.~McKilliam, Barry~G.~Quinn, I.~Vaughan~L.~Clarkson, Bill~Moran and Badri~N.~Vellambi%
    \thanks{%A preliminary version of some of this material is contained in Part 3 of Robby McKilliam's PhD thesis \cite{McKilliam2010thesis}. 
Robby~McKilliam and Badri Vellambi are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.  Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.   Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.  B. Moran is with the Defense Science Institute, University of Melbourne, Vic. 3010, Australia.}}
% The paper headers 
\markboth{Polynomial phase estimation by least squares phase unwrapping}{DRAFT \today}

@@

One drawback of the phase unwrapping approach is that it does not appear to generalise well to the problem of estimating the parameters in a signal that contains the sum of multiple polynomial phase signals.  This appears to be a negative feature of all polynomial phase estimators based on phase unwrapping.  However, the LSU estimator does generalise in other ways.  For example, the LSU estimator can be extended to deal with polynomial phase signals in more dimensions than one.  Our algorithms and analysis would carry over to this case with only minor modification.  Even more generally, the ideas behind the LSU estimator might apply to other types of phase unwrapping problems such as observed in synthetic aperture radar and magnetic resonance imaging~\cite{Katkovnik_phasela_2009, Nico_phaseunwrappingSAR_2000,Friedlander_PD_phaseunwrapping_1996}. This is beyond the scope of our present paper, but may be the subject of future investigation.

@@

A \term{lattice},  $\Lambda$, is a discrete subset of points in $\reals^n$ for which
\[
   \Lambda = \{\Bbf\ubf \mid \ubf \in \ints^d \}
\]
where $\Bbf \in \reals^{n \times d}$ has rank $d$ and is called the generator matrix.  If $n = d$ the lattice is said to be of full rank.  Lattices are discrete Abelian groups under vector addition.  They are subgroups of the Euclidean group $\reals^n$.  Lattices naturally give rise to tessellations of $\reals^n$ by the specification of a set of coset representatives for the quotient $\reals^n / \Lambda$.  One choice for a set of coset representatives is a fundamental parallelepiped, the parallelepiped generated by the columns of a generator matrix.  Another choice is based on the Voronoi cell, those points from $\reals^n$ nearest (with respect to the Euclidean norm in this paper) to the lattice point at the origin.  It is always possible to construct a rectangular set of representatives, as the next proposition will show.  These rectangular regions are used to describe the aliasing properties of polynomial phase signals in Section~\ref{sec:ident_aliasing}.  These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator.
% in Section~\ref{sec:least-squar-unwr}.

@@

Given a polynomial $g(x) = a_0 + a_1x + \dots + a_m x^m$, let
\[
\coef(g) = \left[ \begin{array}{ccccc} a_0 & a_1 & a_2 & \dots & a_m \end{array} \right]^\prime,
\]
denote the column vector of length $m+1$ containing the coefficients of $g$, where $^\prime$ indicates transpose.  If $y$ and $z$ differ by a polynomial in $\mathcal{Z}$ then we can write $y = z + p$ where $p \in \mathcal{Z}$ and then also $\coef(y) = \coef(z) + \coef(p)$.
%\footnote{In group theory terminology $\coef(\cdot)$ coupled with vector addition is called a \term{group homomorphism}.\index{group homomorphism}}. 
  Consider the set
\[
L_{m+1} = \{ \coef(p) \mid p \in \mathcal{Z} \}
\]
containing the coefficient vectors corresponding to the polynomials in $\mathcal{Z}$.  Since the integer valued polynomials form a basis for $\mathcal{Z}$,
\begin{align*}
L_{m+1} &= \{ \coef(c_0 p_0 + c_1p_1 + \dots + c_mp_m) \mid c_i \in \ints \} \\
&= \{ c_0 \coef(p_0) + \dots + c_m\coef(p_m) \mid c_i \in \ints \}.
\end{align*}
Let
\[
\Pbf = \left[ \begin{array}{cccc} \coef(p_0)& \coef(p_1)& \dots& \coef(p_m)  \end{array} \right]
\]
be the $m+1$ by $m+1$ matrix with columns given by the coefficients of the integer valued polynomials.  Then,
\[
L_{m+1} = \{ \Pbf\ubf \mid \ubf \in \ints^{m+1} \}
\]
and it is clear that $L_{m+1}$ is an $m+1$ dimensional lattice.  That is, the set of coefficients of the polynomials from $\mathcal{Z}$ forms a lattice with generator matrix $\Pbf$. We can restate Lemma~\ref{lem:circpolysampledthm} as:
\begin{corollary}\label{cor:circpolysampledcoef}
Two polynomial phase signals $s(t) = e^{2\pi j y(t)}$  and $r(t) = e^{2\pi j z(t)}$ are aliases if and only if $\coef(y)$ and $\coef(z)$ differ by a lattice point in $L_{m+1}$.
\end{corollary}

@@

This minimisation problem can be posed as a nearest lattice point problem~\cite{McKilliam2009asilomar_polyest_lattice, McKilliam2010thesis}.  Write $SS$ as
\[
SS(\mubf) = \sum_{n=1}^{N}{ \left( \Theta_{n} - W_n - \sum_{k = 0}^{m}{\mu_k n^k} \right)^2 },
\]
where $W_n = \round{\Theta_{n} - \sum_{k = 0}^{m}{\mu_k n^k} }$ for $n = 1, \dots, N$ are integers called \emph{wrapping variables}.  If we consider $W_1,\dots,W_N$ as nuisance parameters to be estimated then $SS$ can be considered as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$, the minimiser with respect to $W_1,\dots,W_N$ being $SS(\mubf)$.  The LSU estimator can be found by jointly minimising over $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$.   This joint minimisation problem can be solved by computing a nearest point in a lattice.  To see this write $SS$ as a function of both $\mu_0,\dots,\mu_m$ and $W_1,\dots,W_N$, that is,
\begin{equation}\label{eq:SS_in_vector_form_with_u_nuisance}
SS(\mubf, \wbf) = \| \thetabf - \Xbf\mubf - \wbf \|^2,
\end{equation}
where $\|\cdot\|^2$ denotes the squared Euclidean norm, $\Xbf$ is the rectangular Vandermonde matrix
\begin{equation}\label{eq:Xvandermonde}
\Xbf = \left[ \begin{array}{ccccc} 
1 & 1 & 1 & \cdots & 1 \\
1 & 2 & 4 & \cdots & 2^m \\
\vdots & \vdots  & \vdots & \ddots & \vdots \\
1 & N & N^2 & \cdots & N^m \\
\end{array} \right],
\end{equation}
and the column vectors $\wbf = [W_1, \dots, W_N]^\prime$ and $\thetabf = [\Theta_1, \dots, \Theta_N]^\prime$.   For fixed $\wbf$ the minimiser of $SS(\mubf, \wbf)$ with respect to $\mubf$ is
\begin{equation} \label{eq:hat_theta_cal}
\mubf(\wbf) = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime( \thetabf - \wbf) = \Xbf^+( \thetabf - \wbf),
\end{equation}
where $\Xbf^+ = (\Xbf^\prime\Xbf)^{-1}\Xbf^\prime$ is the Moore-Penrose pseudoinverse of $\Xbf$.  We may thus compute $\widehat{\mubf}$ as $\widehat{\mubf} = \mubf(\widehat{\wbf})$, where $\widehat{\wbf}$ is the minimiser of
\[
SS(\wbf) = SS(\mubf(\wbf),\wbf) = \| \Qbf\thetabf - \Qbf\wbf \|^2,
\]
where $\Qbf = \Ibf - \Xbf\Xbf^+$ and $\Ibf$ is the $N \times N$ identity matrix\footnote{We have slightly abused notation here by reusing $SS$. This should not cause any confusion as $SS(\mubf)$ and $SS(\mubf,\wbf)$ and $SS(\wbf)$ are easily told apart by their inputs.}.

@@

The theorem makes statements about the \emph{dealiased} difference $\dealias(\tilde{\mubf} - \widehat{\mubf})$ between the true coefficients $\tilde{\mubf}$ and the estimated coefficients $\widehat{\mubf}$ rather than directly about the difference $\tilde{\mubf} - \widehat{\mubf}$.   To see why this makes sense, consider the case when $m=0$, $\tilde{\mu}_0 = -0.5$ and $\widehat{\mu}_0 = 0.49$, so that $\tilde{\mu}_0 - \widehat{\mu}_0 = -0.99$.  However, the two phases are obviously close, since the phases $\pm 0.5$ are actually the same.  In this case 
\[
\dealias(\tilde{\mu}_0 - \widehat{\mu}_0) = \fracpart{\tilde{\mu}_0 - \widehat{\mu}_0} = 0.01.
\] 
The same reasoning holds for $m > 0$.

@@

The proof of asymptotic normality places requirements on the pdf $f$ of the phase noise.  The requirement that $\Phi_1, \dots, \Phi_N$ have zero intrinsic mean implies that $f(-\nicefrac{1}{2}) \leq 1$~\cite[Lemma~1]{McKilliam_mean_dir_est_sq_arc_length2010}, so the only case not handled is when equality holds, i.e., when $f(-\nicefrac{1}{2}) = 1$ or when $f(\fracpart{x})$ is discontinuous at $x = -\nicefrac{1}{2}$. In these exceptional cases other expressions for the asymptotic variance can be found (similar to \cite[Theorem 3.1]{Hotz_circle_means_2011}), but this comes at a substantial increase in complexity and for this reason we have omitted them. %Distributions that do not satisfy these requirements are unlikely to be needed in practice.

@@

Tables~\ref{tablm1},~\ref{tablm3} and~\ref{tablm5} show the values of SNR at which the estimators reach a threshold for polynomial phase signals of order $m=1,3$ and $5$.  The values in the tables are approximated using Monte-Carlo simulation with $X_1, \dots, X_N$ pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian.  When $m=3$ the true coefficients are as in Figure~\ref{plot:m3small} and when $m=5$ the true coefficients are as in Figure~\ref{plot:m5} so that the coefficients are inside the range acceptable for the HAF, CPF and CPF-HAF estimators.  When $m=1$ (single frequency estimation) the LSU estimator can be computed exactly using the algorithm from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  In this case the value at which the LSU estimator reaches a threshold is similar to the maximum likelihood periodogram estimator~\cite{Quinn2001,Hannan1973,Rife1974}.  This property was observed in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  Both the LSU and periodogram estimators are strongly consistent, and so, the threshold values can be made arbitrarily small by choosing $N$ large.  This is indicated in the bottom row of Table~\ref{tablm1} by the $-\infty$ symbols.  When $m>1$ the LSU estimator can be exactly computed using the sphere decoder~\cite{Agrell2002,Viterbo_sphere_decoder_1999} in a reasonable amount of time when $N < 60$.  When $N \geq 60$ we use the approximate $K$-best algorithm.  We observe that the $K$-best algorithm produces estimates with the same threshold as the exactly computed LSU estimator when $N = 19$ and $49$.  As $N\rightarrow\infty$ the asymptotic threshold of the HAF estimator has been studied in~\cite{Porat_asympt_HAF_DPT_1996} and these values are indicated in the bottom rows of Tables~\ref{tablm3} and~\ref{tablm5}.  By Theorem~\ref{thm:asymp_proof} the LSU estimator is strongly consistent and the threshold can be made as small as desired by choosing $N$ large.  Whether the LSU estimator computed approximately using the $K$-best algorithm is strongly consistent is an open theoretical question.  That the $K$-best algorithm provides a close approximation of the exact LSU estimator for small $N$ is an encouraging observation.

\begin{table}[t]
\centering
\begin{tabular}{lrr}
$N$ & LSU & ML  \\ \toprule
$16$ & 5 & 2  \\ 
$64$ & -2 & -4 \\ 
$256$ & -6 & -8 \\ 
$1024$ & -14 & -14 \\ 
$\infty$ & $-\infty$ & $-\infty$  \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=1$}
\label{tablm1}
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{lrrrrr}
$N$ & LSU & LSU ($K$-best) & CPF & HAF & ZW \\ \toprule
$19$ & 10 & 10 & 8 & 11 & 31 \\ 
$49$ & 7 & 7 & 5 & 9 & 29 \\ 
$99$ & & 6 & 3 & 7 & 29 \\ 
$199$ & & 3 & 1 & 5 & 28 \\ 
$299$ & & 1 & -1 & 5 & 27 \\ 
$\infty$ & $-\infty$ & & & 3.77 & \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=3$}
\label{tablm3}
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{lrrrrr}
$N$ & LSU & LSU ($K$-best) & CPF-HAF & HAF & ZW \\ \toprule
$19$ & 22  & 22 & 17 & 23 & $>100$ \\ 
$49$ & 17 & 17 & 14 & 21 & $>100$  \\ 
$99$ &  & 15 &  12 &  19 & $>100$ \\ 
$199$ & & 12 & 10 & 18 & $>100$ \\ 
$299$ &  & 11 & 9 & 17 & $>100$ \\ 
$\infty$ & $-\infty$ &  &  & 16.6 &  \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=5$}
\label{tablm5}
\end{table}

@@

We have now shown that the highest order coefficient $\lambda_{m,N}$ converges as required. The remaining coefficients will be shown to converge by induction.  Assume that $N^k \lambda_{k,N} \rightarrow 0$ for all $k=r+1, r+2, \dots, m$.  That is, assume that the $m-r$ highest order coefficients all converge as required by the lemma. Let
\[
z_{N,r}(n) = \lambda_{0,N} + \lambda_{1,N} n + \dots + \lambda_{r,N} n^r.
\]
Because the $m-r$ highest order coefficients converge we can write $z_N(n) = z_{N,r}(n) + \gamma_N(n)$ where 
\[
\sup_{n\in\{1,\dots,N\}}\abs{\gamma_N(n)} \rightarrow 0 \qquad \text{as $N\rightarrow\infty$}. 
\]
Now the bound from \eqref{eq:Deltazfracbound}, but applied using the $r$th difference, gives
 \begin{equation}\label{eq:zrbound}
\begin{split}
\left|\fracpart{  \Delta_h^r z_N(n')}\right| &= \left|\fracpart{  \Delta_h^r\gamma_N(n') + \Delta_h^r z_r(n') }\right| 
\\ &= |\fracpart{ \epsilon + h^r r! \lambda_{r,N} }| < 2^r\delta,
\end{split}
 \end{equation}
 where
\[
\epsilon = \Delta_h^r \gamma_N(n') \leq 2^r \sup_{n\in\{1,\dots,N\}}\abs{\gamma_N(n)} \rightarrow 0
\] 
as $N\rightarrow\infty$.  Choose $\delta$ and $\epsilon$ such that $2^r\delta < \tfrac{1}{4}$ and $|\epsilon| < \tfrac{1}{4}$.  Then, from \eqref{eq:zrbound} and from Lemma~\ref{lem:fracpartinternalsumlessdelta},
\[
\abs{\fracpart{ h^r r! \lambda_{r,N} }} < 2^r\delta + \abs{\epsilon}
\]
for all $h$ such that $1 \leq h \leq \tfrac{N}{2m}$.  Choosing $2^r\delta + |\epsilon| < 2^{-2r-1}$ and using the same iterative process as for the highest order coefficient $\lambda_{m,N}$  (see~\eqref{eq:startiterativearg}~to~\eqref{eq:enditerativearg}) we find that $N^r \lambda_{r,N} \rightarrow 0$ as $N\rightarrow\infty$.  The proof now follows by induction.
 \end{proof}
